---
title: "HW2"
subtitle: "Summary Report for Mushroom Dataset"
date: today
author: Chun-Ju Chou
format:
 pdf:
    include-in-header:
      - text: |
         \usepackage{setspace,relsize}
         \usepackage{geometry}
         \geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
#mainfont: "Microsoft JhengHei UI"
#mainfont: "Microsoft JhengHei"
mainfont: "Microsoft JhengHei Bold"
toc: true
lang: zh-Tw
documentclass: article
pdf-engine: xelatex
execute:
  tidy: true
  echo: true
  warning: false
  message: false
---

## Data information

This dataset was compiled by Dennis Wagner on 05 September 2020. It includes 173 species of mushrooms with caps from various families and one entry for each species. Each species is identified as definitely edible, definitely poisonous, or of unknown edibility and not recommended (the latter class was combined with the poisonous class). Of the 20 variables, 17 are nominal and 3 are metrical. The values of each nominal variable are a set of possible values and for the metrical variables a range of possible values.

## Variable Definition

| Variable (Data Type) | Definition |
|:-------------------|:---------------------------------------------------|
| family (multinomial) | String of the name of the family of mushroom species |
| name (multinomial) | String of the of the mushroom species |
| class (binary) | poisonous=p, edibile=e |
| cap-diameter (metrical) | float number(s) in cm, two values=min max, one value=mean |
| cap-shape (nominal | bell=b, conical=c, convex=x, flat=f, sunken=s, spherical=p, others=o |
| cap-surface (nominal) | fibrous=i, grooves=g, scaly=y, smooth=s,shiny=h, leathery=l, silky=k, sticky=t, wrinkled=w, fleshy=e |
| cap-color (nominal) | brown=n, buff=b, gray=g, green=r, pink=p, purple=u, red=e, white=w, yellow=y, blue=l, orange=o, black=k |
| does-bruise-bleed (nominal) | bruises-or-bleeding=t,no=f |
| gill-attachment (nominal) | adnate=a, adnexed=x, decurrent=d, free=e, sinuate=s, pores=p, none=f, unknown=? |
| gill-spacing (nominal) | close=c, distant=d, none=f |
| gill-color (nominal) | see cap-color + none=f |
| stem-height (metrical) | float number(s) in cm, two values=min max, one value=mean |
| stem-width (metrical) | =bulbous=b, swollen=s, club=c, cup=u, equal=e, rhizomorphs=z, rooted=r |
| stem-surface (nominal) | see cap-surface + none=f |
| stem-color (nominal) | see cap-color + none=f |
| veil-type (nominal) | partial=p, universal=u |
| veil-color (nominal) | see cap-color + none=f |
| has-ring (nominal) | ring=t, none=f |
| ring-type (nominal) | cobwebby=c, evanescent=e, flaring=r, grooved=g, large=l, pendant=p, sheathing=s, zone=z, scaly=y, movable=m, none=f, unknown=? |
| spore-print-color (nominal) | see cap color |
| habitat (nominal) | grasses=g, leaves=l, meadows=m, paths=p, heaths=h, urban=u, waste=w, woods=d |
| season (nominal) | spring=s, summer=u, autumn=a, winter=w |

## Data Description

```{r}
#| results: asis
# R Interface to Python
library(reticulate)
library(Hmisc)
mushroom <- read.csv("primary_data.csv", sep=';')
latex(describe(mushroom), file="")
```

## Table One

```{r}
library(table1)
table1(~ class, data = mushroom)
```
